---
title: "build some sp objects manually"
author: "Jacob Strunk"
date: "6/5/2020"
output:
  md_document:
    variant: gfm
---

# This example demonstrates creation of some sp objects. 

You can find a more complete demonstration of working with spatial object in R here: https://mgimond.github.io/Spatial/reading-and-writing-spatial-data-in-r.html

**Build SpatialPointsDataframe**

Dealing with points is relatively easy compared to polygons or lines

```{r, message=FALSE, warning=FALSE}

library(raster)
library(sp)
library(rgdal)
library(sf)

# Approach 1
npts = 50
pts_df = data.frame(id=1:npts,x=rnorm(npts)*1000+10e6, y=rnorm(npts)*1000+10e6 , z=rnorm(npts))
coordinates(pts_df) = ~x+y
# Approach 2
npts = 50
pts_df = data.frame(id=1:npts,x=rnorm(npts)*1000+10e6, y=rnorm(npts)*1000+10e6 , z=rnorm(npts))
coordinates(pts_df) = pts_df[,c("x","y")]

pts_df
plot(pts_df)
spplot(pts_df,zcol="z")


```

** Add projection to pts_df and write to a shapefile. **

You will need to figure out your projection string if you don't already have it e.g. you can search here https://www.spatialreference.org/ 

I have selected this projection for this toy example https://www.spatialreference.org/ref/epsg/3692/proj4/

See here for more details on projections in R: https://www.nceas.ucsb.edu/sites/default/files/2020-04/OverviewCoordinateReferenceSystems.pdf
```{r, message=FALSE, warning=FALSE}
some_proj4 = "+proj=lcc +lat_1=47.33333333333334 +lat_2=45.83333333333334 +lat_0=45.33333333333334 +lon_0=-120.5 +x_0=500000.0001016001 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +to_meter=0.3048006096012192 +no_defs"

proj4string(pts_df) = some_proj4

#shapefiles are not great, but they are the most common spatial objects (field name restrictions, file size limitations, proprietary, etc.)
if(file.exists("c:/temp/example_pts.shp")) unlink("c:/temp/example_pts.shp",force=T)
rgdal::writeOGR(pts_df,dsn="c:/temp","example_pts1", driver = "ESRI Shapefile", overwrite_layer = T)

#A better option is the open source geopackage (it's like an ESRI file geodatabase). Unfortunately rgdal has pretty lousy support for geopackages as far as I can tell, e.g it cannot append. 
if(file.exists("c:/temp/demo_geopackage.gpkg")) try(unlink("c:/temp/demo_geopackage.gpkg",force=T))
rgdal::writeOGR(pts_df,dsn="c:/temp/demo_geopackage.gpkg","example_pts1", driver = "GPKG")

#the sf package does a better job, but you have to convert from sp to sf first... It seems like sf objects are a better data model (basically an R data frame), but most spatial packages in R still use sp objects.
pts_df_sf = sf::st_as_sf(pts_df)

#this add a new layer to the geopackage - not possible with rgdal (for geopackages)
if(file.exists("c:/temp/demo_geopackage1.gpkg")) try(unlink("c:/temp/demo_geopackage1.gpkg",force=T))
sf::st_write(pts_df_sf,dsn="c:/temp/demo_geopackage1.gpkg", "example_pts2", driver="GPKG") 

#we can even append new rows to an existing object using sf - definitely not possible with rgdal (for geopackages)
pts_df_sf2 = pts_df_sf  
row.names(pts_df_sf2 ) = as.numeric(row.names(pts_df_sf )) + max(as.numeric(row.names(pts_df_sf ))) 
pts_df_sf2$id = pts_df_sf$id + max(pts_df_sf$id)
sf::st_write(pts_df_sf2,dsn="c:/temp/demo_geopackage1.gpkg", "example_pts2", driver="GPKG",append=T) 

#verify that append works
pts_df_sf3 = sf::read_sf(dsn="c:/temp/demo_geopackage1.gpkg", "example_pts2")

#notice that pts_df_sf3 has npts*2 records ...
nrow(pts_df_sf)
nrow(pts_df_sf2)
nrow(pts_df_sf3)

knitr::kable(tail(pts_df_sf3,10))
#DT::datatable(pts_df_sf3)
```


** buffer points to create polygons**
this is an easy way to make a SpatialPolygonsDataFrame
```{r, message=FALSE, warning=FALSE}

bf1 = raster::buffer(pts_df , dissolve=FALSE)
bf1

```

**Create SpatialPolygonsDataFrame**

Making a SpatialPolygonsDataFrame manually is a lot of work

 1. create single polygon objects sp::Polygon(coords, hole=as.logical(NA))
 2. make a list of polygon objects and combine polygons sp::Polygons(srl, ID)
 3. combine list of Polygons  sp::SpatialPolygons(Srl, pO, proj4string=CRS(as.character(NA)))
 4. finally, merge spatial data with data frame sp::SpatialPolygonsDataFrame(Sr, data, match.ID = TRUE)

now lets make a circle
```{r}

  #number of vertices
  nvtx = 100
  #split circle into nvtx slices - add zero to end to close polygon
  rads0 = c(seq(0,2*pi,length.out = nvtx),0)
  
  #compute x,y coordinates
  radius0 = 100
  xcds = radius0*cos(rads0)
  ycds = radius0*sin(rads0)
  plycds = data.frame(x=xcds, y=ycds)
  
  #verify that we built a circle
  plot(plycds,asp=1,type="l")
  
  #build sp objects
  ply1a = sp::Polygon(plycds)
  ply1b = sp::Polygon(plycds+50)
  ply2a = sp::Polygons(list(ply1a),ID=1)
  ply2b = sp::Polygons(list(ply1b),ID=2)
  ply2c = sp::Polygons(list(ply1a,ply1b),ID=1)
  #same data, two ways to store it
  ply3a = SpatialPolygons(list(ply2a,ply2b))
  ply3b = SpatialPolygons(list(ply2c))
  #take a look
  ply3a
  plot(ply3a)
  ply3b
  plot(ply3b)
  #finish the object
  ply4a = SpatialPolygonsDataFrame( ply3a , data.frame(id=1:length(ply3a)))
  ply4b = SpatialPolygonsDataFrame( ply3b , data.frame(id=1:length(ply3b)))
  #look at final products
  ply4a
  ply4b
  par(mfrow=c(1,2))
  plot(ply4a,col=ply4a@data[,"id"])
  plot(ply4b,col=ply4b@data[,"id"])
  #pretty interesting, in the second approach, R interprets the overlap as void...
  
```

**Create SpatialLinesDataFrame**

Lines use a similar process to polygons. I have taken a shortcut here and based my example on the sp vignette.

```{r}

  # based on example from the sp vignette:
  l1 = cbind(c(1,2,3),c(3,2,2))
  rownames(l1) = letters[1:3]
  l1a = cbind(l1[,1]+.05,l1[,2]+.05)
  rownames(l1a) = letters[1:3]
  l2 = cbind(c(1,2,3),c(1,1.5,1))
  rownames(l2) = letters[1:3]
  Sl1 = Line(l1)
  Sl1a = Line(l1a)
  Sl2 = Line(l2)
  S1 = Lines(list(Sl1, Sl1a), ID="a")
  S2 = Lines(list(Sl2), ID="b")
  Sl = SpatialLines(list(S1,S2))
  summary(Sl)
  plot(Sl, col = c("red", "blue"))
  
  #create sp dataframe
  sldf = SpatialLinesDataFrame(Sl, data.frame(id = names(Sl), someY=rnorm(length(Sl)),row.names=names(Sl)), match.ID = TRUE)
  
  #plot spatial objects
  sp::spplot(sldf,"someY",lwd=3)

```

buffer lines to make polygons
```{r}

  lnbuff1 = raster::buffer(sldf,.5,dissolve=F) 
  lnbuff2 = raster::buffer(sldf,.5,dissolve=T) 
  #or
  lnbuff1 = rgeos::gBuffer(sldf,width=.5,byid = F ) 
  lnbuff2 = rgeos::gBuffer(sldf,width=.5,byid = T) 
  
  #look at results
  lnbuff1
  lnbuff2
  par(mfrow=c(1,2))
  raster::plot(lnbuff1)
  raster::plot(sldf,col=1:length(sldf),add=T)
  raster::plot(lnbuff2)
  raster::plot(sldf,col=1:length(sldf),add=T)

```
