---
title: "Example_workflow"
author: "Jacob Strunk"
date: "5/7/2020"
output: html_document
---

load packages - requires devtools package
```{r , eval = FALSE}
#load required packages from github
# lidR and rlas can be a bit tricky to install
# if you are strugglign to install from github, look at the console and see which dependency is 
# causing the failure. For example the "glue" package was failing when I prepared this workflow.
# To overcome the issue I used install.packages("glue") and then tried install_github("Jean-Romain/lidR")
# you may also need to compile something from source if the wrong package version is available, e.g.
# install.packages("glue", type="source") which also fails for some packages...

# you can optionally also install from binaries with install.packages(c("rlas","lidR"))


devtools::install_github("Jean-Romain/rlas")
devtools::install_github("Jean-Romain/lidR")
devtools::install_github("jacobstrunk001/RSForInvt")

```

#convert dtm raster to fusion .dtm rasters
```{r}

RSForInvt::raster2FUSION(
  "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\DTM\\"
  ,"D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\DTM_fusion\\"
  )


```


#scan a lidar project
```{r}

 #scan las/laz files and build vector files in a .gpkg (geopackage)
 las_info = RSForInvt::scan_las(
    project="TN_lidar"
    ,project_year="2018"
    ,proj4_name=NA #"NAD_1983_HARN_StatePlane_Washington_South_FIPS_4601_Feet"
    ,proj4= NA 
    ,dir_las = "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\lidar_tiles\\"
    ,pattern="[.]la.$"
    ,notes=""
    ,create_polys=T
    ,recursive = T
    ,return=T
    )

#look at results that are returned directly when return=T
  las_info$project_id
  View(las_info$las_ids) 
  plot(las_info$plys) 

#grab the exact same data from geopackage
  plys_las = rgdal::readOGR(dsn = "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\lidar_tiles\\manage_las\\manage_las.gpkg" , layer = "las_polys" )
  con_gpkg = dbConnect(RSQLite::SQLite(), "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\lidar_tiles\\manage_las\\manage_las.gpkg")
  las_ids = dbReadTable(con_gpkg,"las_id")
  proj_id = dbReadTable(con_gpkg,"project_id")  
  dbDisconnect(con_gpkg)
  
  plot(plys_las)
  proj_id
  View(las_ids)
  
  

```


#scan dtms
```{r}
#scan dtms and build vector files in a .gpkg (geopackage)
 dtm_info = RSForInvt::scan_dtm(
    project="TN_lidar"
    ,project_year="2018"
    ,proj4_name=NA #"NAD_1983_HARN_StatePlane_Washington_South_FIPS_4601_Feet"
    ,proj4= NA 
    ,dir_dtm = "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\DTM_fusion"
    ,pattern="[.]dtm$"
    ,notes=""
    ,create_polys=T
    ,recursive = T
    ,return=T
    )

#look at results that are returned directly 
  dtm_info$project_id
  View(dtm_info$dtm_ids) 
  plot(dtm_info$plys) 

#grab the exact same data from geopackage
  plys_dtm = rgdal::readOGR(dsn = "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\DTM_fusion\\manage_dtm\\manage_dtm.gpkg" , layer = "dtm_polys" )
  con_gpkg_dtm = dbConnect(RSQLite::SQLite(), "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\DTM_fusion\\manage_dtm\\manage_dtm.gpkg")
  dtm_ids = dbReadTable(con_gpkg_dtm,"dtm_id")
  proj_id_dtm = dbReadTable(con_gpkg_dtm,"project_id")  
  
  proj_id_dtm
  View(dtm_ids)
  plot(plys_dtm)
  
  dbDisconnect(con_gpkg)

if(F){
  
  plot(subset(plys_dtm,subset = 1:nrow(plys_dtm) ==1  ))
  plot( plys_las , add=T , border = "red")
    
}

```


#build project
```{r }

#build project
# !!A really important consideration is the tile size!!
proj_tn = RSForInvt::project_create(
  dir_las="D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\lidar_tiles\\"
  ,dir_dtm="D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\Data\\Tennessee\\DTM_fusion\\"
  ,path_gpkg_out="D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\project\\TNLidar_RSForInvtProject.gpkg"
  ,layer_project = "RSForInvt_prj"
  ,layer_config = "RSForInvt_config"
  ,overwrite_project = T
  ,project_dtm="lidar_dtm"
  ,project_las="lidar_las"
  ,dtm_year="2018"
  ,las_year="2018"
  ,do_scan_dtms=F #we already scanned the dtm folder - ok, to rescan, but slower
  ,do_scan_las=F #we already scanned the las folder - ok, to rescan, but slower
  ,tile_size=6000
  ,pixel_size=66
  #,xmn=561066,xmx=2805066,ymn=33066,ymx=1551066
  ,proj4 = "+proj=lcc +lat_1=47.33333333333334 +lat_2=45.83333333333334 +lat_0=45.33333333333334 +lon_0=-120.5 +x_0=500000.0001016001 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=us-ft +no_defs"
  ,mask=NA
  ,return=T
)

#look at project components
  proj_tn$las_tiles_bfr
  proj_tn$dtm_tiles_bfr
  proj_tn$project_plys

```


#create gridmetrics output csvs
```{r }

  #use project tiles, iterate across project, and run FUSIOn gridmetrics.exe for each processing tile
  #this is essentially what areaprocesser does...
  gmi=RSForInvt::run_gridmetrics(
    proj_gpkg_path="D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\project\\TNLidar_RSForInvtProject.gpkg"
    ,dir_out="D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\"
    ,n_core=4
   )



```


#load csvs into sqlite database
```{r }

  #load csvs into sqlite database
  RSForInvt::csv_to_sqlite(
      db="D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\gridmetrics_sqlite\\gridmetrics.sqlite"
       ,csv_folder = "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\gridmetrics_csv"
       ,tb_summary="gm_summary"
       ,tb_csv="gm"
       ,project="TN_2018"
       ,resolution="66"
       ,units="feet"
       ,proj4="+proj=lcc +lat_1=47.33333333333334 +lat_2=45.83333333333334 +lat_0=45.33333333333334 +lon_0=-120.5 +x_0=500000.0001016001 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=us-ft +no_defs"
       ,notes=""
       ,skip_loaded=T
       ,n_load=NA
       ,use_col_classes=T
       )



```


#create rasters from sqlite database
```{r }
  library(DBI)
  library(RSQLite)

  proj4_this = "+proj=lcc +lat_1=47.33333333333334 +lat_2=45.83333333333334 +lat_0=45.33333333333334 +lon_0=-120.5 +x_0=500000.0001016001 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=us-ft +no_defs"

  #take all fields in sqlite and create a raster
  #for each field, or provide a subset. colsSomeX is
  # a subset that I prepared...
  db_gridmetrics = dbConnect(RSQLite::SQLite(), "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\gridmetrics_sqlite\\gridmetrics.sqlite")
    
  RSForInvt::sqlite_to_raster(
    db=db_gridmetrics
    ,tb_csv="gm"
    ,colsxy = c("center_x","center_y")
    ,cols2Raster = colsSomeX()
    ,format = ".img"
    ,dirOut = "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\gridmetrics_rasters"
    ,crs = proj4_this
    ,nProc = 5
    ,doDebug=F
    ,doBuild=F
  
  )

  dbDisconnect(db_gridmetrics)

```

#a single raster is inefficient for this dataset. The data is composed of a sample of lidar tiles
#spread across the entire state. Instead we will see how to create rasters from each lidar tile
#and then glue them together using the gdal .vrt file type (accessible in R with gdalUtils package).
#first we create views with subsets of the data and convert them to rasters.
```{r }
  library(DBI)
  library(RSQLite)
  library(gdalUtils)

  proj4_this = "+proj=lcc +lat_1=47.33333333333334 +lat_2=45.83333333333334 +lat_0=45.33333333333334 +lon_0=-120.5 +x_0=500000.0001016001 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=us-ft +no_defs"

  #take all fields in sqlite and create a raster
  #for each field, or provide a subset. colsSomeX is
  # a subset that I prepared...
  db_gridmetrics = dbConnect(RSQLite::SQLite(), "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\gridmetrics_sqlite\\gridmetrics.sqlite")
    
  #see if there is an indication of tile in gridmetrics data
  dbGetQuery(db_gridmetrics, "select * from gm limit 5")
  
  #lets use identifier to split data
  ids = unlist(dbGetQuery(db_gridmetrics, "select distinct identifier from gm"))
  
  #create views and rasters
  for(i in 1:length(ids)){
    
    vw_i = paste("vw_gm_id_",ids[i],sep="")
    qry_vw_id = dbSendQuery(db_gridmetrics, paste("create view if not exists",vw_i,"as select * from gm where identifier = ",ids[i]))
  
    RSForInvt::sqlite_to_raster(
      db=db_gridmetrics
      ,tb_gm=vw_i
      ,colsxy = c("center_x","center_y")
      ,cols2Raster = RSForInvt::colsFewX("6_00")
      ,format = ".img"
      ,dirOut = "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\gridmetrics_rasters\\TN"
      ,raster_prefix = paste("TN_id_XXXX_",ids[i],"_",sep="")
      ,crs = proj4_this
      ,nProc = 5
      ,doDebug=F
      ,doBuild=F
    )

  }  
  dbDisconnect(db_gridmetrics)
  
  #create .vrt files - a virtual raster format that can "point" to existing rasters
  gm_nms = colsFewX("6_00")
  for(i in 1:length(gm_nms)){

    rasters_i = list.files("D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\gridmetrics_rasters\\TN"
                           ,pattern = paste(".*",gm_nms[i],".*img$",sep=""),full.names=T)
    vrt_i = paste("D:/Box/VMARS/Projects/DAP_evaluation_Meston/R/DAP_Lidar_analysis/RSForInvt/gridmetrics_rasters/TN_VRT/",gm_nms[i],".vrt",sep="")
    gdalUtils::gdalbuildvrt(
      rasters_i
      , vrt_i
    )

  }
  
```


#create .dtm canopy models
```{r }

  gmi=RSForInvt::run_canopyModel(
    proj_gpkg_path="D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\project\\TNLidar_RSForInvtProject.gpkg"
    ,dir_out="D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\"
    ,n_core=4
   )



```

#(optional) convert .dtm canopy models to .img and build composite .vrt files
```{r }

  paths_dtms = list.files("D:/Box/VMARS/Projects/DAP_evaluation_Meston/R/DAP_Lidar_analysis/RSForInvt/canopymodel/TN/individual_dtm"
                          ,pattern="[.]dtm$",full.names=T)
  paths_imgs = file.path("D:/Box/VMARS/Projects/DAP_evaluation_Meston/R/DAP_Lidar_analysis/RSForInvt/canopymodel/TN/individual_img"
                         ,gsub("[.]dtm$",".img",basename(paths_dtms))
  )
  
  for(i in 1:length(paths_dtms)){
    
    dtm_i = RSForInvt::read_dtm(paths_dtms[i])
    raster::writeRaster(dtm_i,paths_imgs[i])
    
  }

  paths_imgs1 = list.files("D:/Box/VMARS/Projects/DAP_evaluation_Meston/R/DAP_Lidar_analysis/RSForInvt/canopymodel/TN/individual_img"
                         ,pattern="[.]img$",full.names=T)
  
  gdalUtils::gdalbuildvrt(
      paths_imgs1
      , "D:/Box/VMARS/Projects/DAP_evaluation_Meston/R/DAP_Lidar_analysis/RSForInvt/canopymodel/TN/merged_vrt/tn_dtms.vrt"
    )

```

#merge .dtm canopy models and create ITD objects - FUSION
```{r }

csms = list.files("D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\canopymodel\\individual" ,pattern="[.]dtm" , full.names=T)
writeLines(csms, "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\canopymodel\\individual\\csm_list.txt")
shell("C:\\FUSION\\mergedtm.exe /disk D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\canopymodel\\merged\\tn_dsm.dtm D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\canopymodel\\individual\\csm_list.txt",wait=F)

#fails in this example because extent is too large
shell("c:\\fusion\\treeseg.exe /shape D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\canopymodel\\merged\\tn_dsm.dtm 30 D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\ITD_merged\\tn_ITD.csv",wait=T)


```


#build canopy models using lidR
```{r }
if(F){
  install.packages("BiocManager")
  BiocManager::install("EBImage" , update = F)
  install.packages("ForestTools")
}
library(EBImage)
install.packages("EBImage",type="source")
csm = raster::raster("D:/Box/VMARS/Projects/DAP_evaluation_Meston/R/DAP_Lidar_analysis/RSForInvt/canopymodel/TN/merged_vrt/tn_dtms.vrt")
ws1 = lidR::watershed(csm, th = 30)()


library("ForestTools")
library(raster)

csm = raster::raster("D:/Box/VMARS/Projects/DAP_evaluation_Meston/R/DAP_Lidar_analysis/RSForInvt/canopymodel/TN/merged_vrt/tn_dtms.vrt")
lin <- function(x){x * 0.05 + 0.6}
ttops <- vwf(CHM = csm, winFun = lin, minHeight = 30)
crowns <- mcws(treetops = ttops, CHM = csm , minHeight = 6, verbose = FALSE , OSGeoPath = 'C:\\OSGeo4W64')

```



#segment individual trees for each tile
```{r }

 RSForInvt::run_treeseg(
    dir_dtms = "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\canopymodel\\individual"
    ,dtm_paths = NA
    ,dir_out = "D:\\Box\\VMARS\\Projects\\DAP_evaluation_Meston\\R\\DAP_Lidar_analysis\\RSForInvt\\ITD"
    ,switches = "/shape"
    ,treeseg_path = "c:\\fusion\\treeseg.exe"
    ,ncore = 4
  )


```


#clip plots from lidar
```{r }

#load lidar exents



#generate pseudo plot locations



#create FIA plots at sample points



#clip plots




```


#summarize plot clips
```{r }



```


#interpolate raster data to plot locations (alternative to plot clips)
```{r }



```

